---
output:
  html_document:
    theme: flatly
    css: styles.css
---

<link rel="stylesheet" href="styles.css" type="text/css" />
<link rel="stylesheet" href="academicons/css/academicons.min.css"/>


## Likelihood Function

<br>
<br>

#### **Likelihood Function**
<br>

Denote with $p(x|\theta)$, the PMF(probability mass function) of each $X_i$,
$L(\theta)=\displaystyle\prod_{i=1}^{n}p(x_i|\theta)$

<br>

#### **Log-Likelihood Function**
$l(\theta)=\displaystyle\sum_{i=1}^{n}logp(x_i|\theta)$

<br>

#### **Likelihood Function of Binomial Sample**
$L(\pi)={n\choose k}\pi^y(1-\pi)^{n-y}$, y=0,1,...,n
<br>
Written with R function
<br>

```{r}
llfun <- function(pi,n,observation){
  choose(n,observation)*(pi^observation)*(1-pi)^(n-observation)
}
```

<br>

#### **Likelihood Function of Binomial Sample**
$l(\pi)=ln{n\choose k}+yln\pi+(n-y)ln(1-\pi)$, y=0,1,...,n
<br>
Written with R function
<br>

```{r}
loglik <- function(pi,n,observation){
  log(choose(n,observation))+(observation*log(pi))+(n-observation)*log(1-pi)
}
```

<br>

#### **The maximum likelikhood estimator is the parameter value that maximizes the likelihood function**
<br>
The demonstrates the value of $\pi$ that maximizes the log-likelihood is $\frac{y}{n}$

<br>

#### **Examples**
OZQ Hospital is interested in determining its rate of deliveries between the hours of midnight and 8 am.They randomly select 10 nights from last year and observe the following number of deliveries per each night:3, 3, 2, 4, 3, 1, 0, 3, 5, 1. The hospital knows that the distribution for the number of deliveries between thehours of midnight and 8 a.m. is a Poisson distribution. They want to find the maximum likelhood estimate of $\lambda$ , the rate of deliveries per each midnight to 8 am shift.

<br>

Using $L(\theta)=\displaystyle\prod_{i=1}^{n}p(x_i|\theta)$, and $p(x_i|\lambda)=\displaystyle\prod_{i=1}^{n}\frac{\lambda^{x_i}e^{-\lambda}}{x_i!}$ to plot a likelihood function

<br>

```{r}
likelihood <-function(lambda, x) {
pdf_xi =vector(length =length(x))

for(i in 1:length(x)){
  pdf_xi[i] = ((lambda^x[i])*(exp(-lambda)))/factorial(x[i])
  }
prod(pdf_xi)
}
x =c(3, 3, 2, 4, 3, 1, 0, 3, 5, 1)
lambdas <-seq(0.01, 5.5, by=0.01)
likelihoods <-sapply(lambdas,likelihood,x = x)
plot(lambdas, likelihoods, type = "l", lwd=4, 
     xlab = "Lambda values", 
     ylab = "Likelihood",
     main ="Likelihood function")
abline(v = lambdas[which(likelihoods== max(likelihoods))], 
       col="red", lwd=3, lty=2)
```

<br>

Using $l(\theta)=\displaystyle\sum_{i=1}^{n}logp(x_i|\theta)=log(L(\theta))$, to plot a log-likelihood function

<br>

```{r}
log_likelihood <-function(lambda, x) {
  log(likelihood(lambda, x))
  }
#Using x and lambdas defined above
log_likelihoods <-sapply(lambdas,log_likelihood,x = x)
plot(lambdas, log_likelihoods, type = "l", lwd=4, 
     xlab = "Lambda values", ylab = "Log-likelihood",
     main ="Log-Likelihood function")
abline(v = lambdas[which(likelihoods== max(likelihoods))], col="red", lwd=3, lty=2)
```